A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{hyperref}
\newcommand{\missingNumber}{\textcolor{red}{XX}\xspace}
\newcommand{\missingPercentage}{\textcolor{red}{XX\%}\xspace}
\newcommand{\missingTable}{\textcolor{red}{XXTable}\xspace}
\newcommand{\missingGraph}{\textcolor{red}{XXGraph}\xspace}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{paralist}
\usepackage{xspace}
\usepackage{url}

@@

In this section we turn our attention to the expression passed for evaluation to
\eval. \eval accepts all types of R objects as inputs. In our corpus, we observe
that \AllValueInputEvalCallPerc of all inputs to \eval are values, i.e., they
evaluate to themselves. The remaining \AllExpressionInputEvalCallPerc inputs are
expressions which are reduced to values by \eval. % Furthermore, the proportion
of values % passed to core \eval call is \CoreEvalValueInputPerc compared to a
significantly % higher \PackageEvalValueInputPerc for package \eval calls. To
investigate this, we look at the type of expressions accepted by \eval callsites
in our corpus. We find that only \AllValueInputEvalSitePerc receive only values
as inputs, \AllExpressionInputEvalSitePerc call sites exclusively receive
expressions as inputs and a very large proportion,
\AllPolymorphicInputEvalSitePerc, receive both expressions and values as inputs
across all program runs. We inspected the eval callsites that accept only values
or both values and expression as inputs. The top three of these call sites
contribute to over ~80\% of all calls to evals. The first one is inside
\c{match.arg} from core. Given a vector of names as strings, this function finds
the default values for parameters with those names in the caller function. Since
the default value for a function parameter can be a complex expression or a
benign value such as \c{NULL}, \c{match.arg} evaluates it through a call to
\eval. This makes the \eval inside \c{match.arg} ``polymorhpic''. The second
such \eval call originates from the popular \ggplot package of R which is a DSL
for plotting. \ggplot has it's own object system, \ggproto, which uses \eval
inside the \ggproto function to determine the super-class from which it
inherits. The default value of the superclass is \c{NULL} but in general could
be an arbitrarily complex expression. The third \eval call is in core R \c{str}
function. This function is used to print a summary of R objects. It internally
maps \eval function on the bindings of an environment which results in inputs of
multiples types being passed to \eval.

@@

From the figure, we observe that majority of inputs to Core \eval functions are
expressions in contrast to package \eval calls which receive mostly environments
as inputs. In fact, \PackageEnvironmentInputEvalCallPerc of \eval calls in
packages receive environments as input expressions. There are only four
callsites in package evals that receive environments as expressions.
\begin{itemize}
\item \c{ggplot2::ggproto} contributes to 99.9\% of package \eval calls that
receive environments as input.
\item \c{R6::generator_funs} function contributes to only 936 \eval calls and
implements the same functionality as \c{ggplot2::ggproto} but for the \emph{R6}
OOP system.
\item \c{future::backtrace} function applies \eval to a future object which is
implemented as an environment. This is called only once.
\item \c{RModel::str.RMmodel} function overloads the core \c{str} method for its
\c{RModel} objects and maps \eval exactly like \c{str}. This is called only
once.
\end{itemize}

>>>>

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{hyperref}
\newcommand{\missingTag}[1]{\textcolor{red}{#1}\xspace}
\newcommand{\missingNumber}{\textcolor{red}{XX}\xspace}
\newcommand{\missingPercentage}{\textcolor{red}{XX\%}\xspace}
\newcommand{\missingTable}{\textcolor{red}{XXTable}\xspace}
\newcommand{\missingGraph}{\textcolor{red}{XXGraph}\xspace}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{paralist}
\usepackage{xspace}
\usepackage{url}

@@

In this section we turn our attention to the expression passed for evaluation to
\eval. The \eval function accepts all types of R objects as inputs, but in the
majority of cases (\AllExpressionInputEvalCallPerc), it is called with
expressions. The remaining \AllValueInputEvalCallPerc are values, \ie, they
evaluate to themselves.
%
%Furthermore, the proportion of values passed to core
%\eval call is \missingTag{CoreEvalValueInputPerc} compared to a significantly higher
%\missingTag{PackageEvalValueInputPerc} for package \eval calls.
%
Passing a value into an eval is effectively an no-op so we look closely at the
expression types the call sites pass to eval across all the runs.
\AllValueInputEvalSitePerc of all call sites always pass values,
\AllExpressionInputEvalSitePerc pass expressions and a very large proportion,
\AllPolymorphicInputEvalSitePerc, pass both expressions and values as inputs
across all program runs. We inspected the eval callsites that accept only values
or both values and expression as inputs. The top three of these call sites
contribute to over ~80\% of all calls to evals. The first one is inside
\c{match.arg} from core. Given a vector of names as strings, this function finds
the default values for parameters with those names in the caller function. Since
the default value for a function parameter can be a complex expression or a
benign value such as \c{NULL}, \c{match.arg} evaluates it through a call to
\eval. This makes the \eval inside \c{match.arg} ``polymorhpic''. The second
such \eval call originates from the popular \ggplot package of R which is a DSL
for plotting. \ggplot has it's own object system, \ggproto, which uses \eval
inside the \ggproto function to determine the super-class from which it
inherits. The default value of the superclass is \c{NULL} but in general could
be an arbitrarily complex expression. The third \eval call is in core R \c{str}
function. This function is used to print a summary of R objects. It internally
maps \eval function on the bindings of an environment which results in inputs of
multiples types being passed to \eval.

@@

From the figure, we observe that majority of inputs to Core \eval functions are
expressions in contrast to package \eval calls which receive mostly environments
as inputs. In fact, \PackageEnvironmentInputEvalCallPerc of \eval calls in
packages receive environments as input expressions. There are only four
callsites in package evals that receive environments as expressions.
\begin{compactitem}[$-$]
\item \c{ggplot2::ggproto} contributes to 99.9\% of package \eval calls that
receive environments as input.
\item \c{R6::generator_funs} function contributes to only 936 \eval calls and
implements the same functionality as \c{ggplot2::ggproto} but for the \emph{R6}
OOP system.
\item \c{future::backtrace} function applies \eval to a future object which is
implemented as an environment. This is called only once.
\item \c{RModel::str.RMmodel} function overloads the core \c{str} method for its
\c{RModel} objects and maps \eval exactly like \c{str}. This is called only
once.
\end{compactitem}
